Predicting instabilities of a tuneable ring laser with an iterative map model

### 2020-11-24 Papers/Journal/Tuneable_Lasers.tex
Fixed table and some notation

\usepackage{subcaption}
\usepackage{tikz}

@@

To proceed we must choose the form of the amplification term, $g(A)$. We assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}. This reduces \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the incoming energy is $E$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse within the Er-doped fibre. Repeating this process for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 L_D\beta_2/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the output coupler (depending on the layout of the laser cavity the loss may instead take the form $L(A) = R \exp( - \alpha L_T )A$), $L_T$ is the total length of the laser circuit, $L_D$ is the length of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the output coupler, and $\mathcal{F}$ denotes the Fourier transform. Finally, we must consider the modulation. The modulation is considered to be applied externally in which ever way the operator sees fit. For simplicity the representation is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not particularly restrictive. Calcaterra and Boldt have shown that Gaussians can form a basis for $L^2(\mathbb{R})$~\cite{calcaterra2008a}. Therefore, we can approximate any modulation function in $L^2(\mathbb{R})$ by a sum of \eqref{eq:modform}.

@@

Revisiting the process maps \eqref{eq:gainform}--\eqref{eq:modform} yields the new, non-dimensional, mappings---after dropping the tildes---
\begin{align}
	G(A) &= \left(E^{-1} W \left( a E \textrm{e}^{E}\right) \right)^{1/2} A, & F(A) &= A \exp( i b |A|^2), \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters, as defined by the values in Table~\ref{tab:values},
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2.
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
Here, $a$ is a measure of the strength of the gain fibre, $h$ is the loss due to absorption and the output coupler, $b$ is the ratio of the frequency shift due to the Kerr nonlinearity to the width of the pulse, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, notice that in this non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:ndparam} have their own independent, non-dimensional parameter.

% Table of parameter values
\begin{table*}[tbp]
	\centering
	\caption{Range of variation of various parameters.}
 	\label{tab:values}
 	\begin{tabular}{lcll}
		\hline\noalign{\smallskip}
		Parameter & Symbol & Value & Sources \\
		\hline\noalign{\smallskip}
		Absorption of Fibre & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  &~\cite{burgoyneemail, shtyrina2017, tomlinson1984, usechak2005} \\
		Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ &~\cite{agrawal2002, agrawal2013, burgoyne2014} \\
		Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ &~\cite{agrawal2013, finot2008, usechak2005} \\
		Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ &~\cite{agrawal2002, agrawal2013, burgoyne2014} \\
		Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ &~\cite{burgoyneemail, tamura1996} \\
		Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ &~\cite{burgoyneemail} \\
		Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ &~\cite{burgoyne2014, shtyrina2017} \\
		Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ &~\cite{bohun2015, burgoyneemail, burgoyne2014} \\
		Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ &~\cite{burgoyneemail, tamura1996, yamashita2009} \\
		Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ &~\cite{burgoyneemail, usechak2005} \\
		Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ &~\cite{burgoyneemail} \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table*}

@@

Consider the pulse
\begin{align}
	A = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \textrm{e}^{i \phi_0},
	\label{eq:A0}
\end{align}
where $P$ is the peak power, $C$ is the chirp, $\sigma^2$ is the variance, and $\phi_0$ is the phase (Fig.~\ref{fig:samplegauss}). By computing $\mathcal{L}(A)$ using \eqref{eq:order}, and equating to $A \textrm{e}^{i \phi}$, we obtain the conditions
\begin{align}
	\frac{\sigma^4}{1 - \sigma^2} &= (\sigma^2 + 2 C s^2)^2 + 4 s^4, \label{eq:sigrelation} \\
	\frac{C}{1 - \sigma^2} &= C + 2 \frac{s^2}{\sigma^2} (1 + C^2), \label{eq:chirprelation} \\
	1 &= \frac{W(a E \textrm{e}^E)}{E} h^2 (1 - \sigma^2)^{1/2} \label{eq:energyrelation}
\end{align}
for \eqref{eq:A0} to be in equilibrium. Manipulating \eqref{eq:sigrelation} and \eqref{eq:chirprelation} we are able to eliminate the chirp, and find an expression in terms of $\sigma$ only, that is,
\begin{align}
	\sigma^8 + 4 s^4 \sigma^6 - 20 s^4 \sigma^4 + 32 s^4 \sigma^2 - 16 s^4 = 0. \label{eq:var}
\end{align}
As \eqref{eq:var} is a quartic in $\sigma^2$, it has an analytic solution, namely,
\begin{equation}
	\begin{split}
		\sigma^2 = \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} & \\
		- s^4 - s^2 & \sqrt{4 + s^4}.
		\label{eq:equilvar}
	\end{split}
\end{equation}
Additionally, we find the chirp to be
\begin{align}
	C = \frac{\sigma^4 \pm \left( \sigma^8 - 16 s^4 \left( 1 - \sigma^2 \right)^2 \right)^{1/2}}{4 s^2 \left( 1 - \sigma^2 \right)},
	\label{eq:chirp}
\end{align}
in terms of $s$ and $\sigma^2$. Moreover, we can obtain the energy of the pulse entering the gain fibre from \eqref{eq:energyrelation}, and thus, the equilibrium energy at the output coupler. Doing so gives
\begin{align}
	E &= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{1 - h^2 \left( 1 - \sigma^2 \right)^{1/2}}.
	\label{eq:analenergy}
\end{align}
By equating the energy of the pulse in \eqref{eq:A0} and \eqref{eq:analenergy}, the peak power is found to be
\begin{align}
	P &= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{\sqrt{\pi} \sigma \left( 1 - h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}.
	\label{eq:analpower}
\end{align}

The nature of \eqref{eq:var}--\eqref{eq:analpower} makes it difficult to work with. However, by asymptotically expanding \eqref{eq:var}--\eqref{eq:analpower} in terms of the dispersion parameter, $s$, we find the simpler relations
\begin{align}
	\sigma^2 &\sim 2s(1 - s) + \bigO{s^3}, \\
	C &\sim 1 - s + \frac{1}{2}s^2 + \bigO{s^3}, \\
	E &\sim \frac{\log (a h^2)}{1 - h^2} - \frac{1}{1 - h^2} \left( 1 + \frac{h^2 \log(a h^2)}{1 - h^2}  \right) s + \bigO{s^2}, \\
	P &\sim \frac{\log(ah^2)}{\sqrt{2 \pi}(1 - h^2)} s^{-1/2} + \bigO{s^{1/2}},
\end{align}
in the case $s \rightarrow 0$. And,
\begin{align}
	\sigma^2 &\sim 1 - \frac{1}{4}s^{-4} + \frac{3}{8}s^{-8} + \bigO{s^{-12}}, \label{eq:siginfty}\\
	C &\sim \frac{1}{2}s^{-2} - \frac{3}{8}s^{-6} + \bigO{s^{-10}},
\end{align}
in the case $s \rightarrow \infty$. We have no asymptotic expansion of the energy or peak power as $s \rightarrow \infty$ since in this limit the gain is unable to balance the energy lost due to modulation, as indicated by the logarithm in \eqref{eq:analenergy}. Expanding the condition $a h^2 (1 - \sigma^2)^{1/2} > 1$, from \eqref{eq:analenergy}, using \eqref{eq:siginfty} (for a first-order approximation), we find 
\begin{align}
	s^* = \left( \frac{a h^2}{2} \right)^{1/2}
\end{align}
is the maximum value of $s$ to facilitate a sustainable pulse.

@@

We now turn our attention to a more quantitative analysis, and characterize the stability of a laser by exploring the parameter space. We focus on the $s$--$b$ plane in particular since $a$ and $h$ effect only the amplitude of the pulse, whereas the interplay of dispersion and nonlinearity give rise to the instabilities. Additionally, as mentioned, a conservative initial energy will effect the dynamics much less than a more intense pulse since the nonlinearity will be negligible while coming to equilibrium. To identify the stability of a pulse we examine the relative change in the pulse's envelope between consecutive round trips of the cavity. We compute this error by
\begin{align}
	\Delta = \frac{\| |A_i| - |A_{i-1}| \|_2}{\| A_{i-1} \|_2},
	\label{eq:error}
\end{align}
where $\| \cdot \|_2$ denotes the $L^2(\mathbb{R})$ norm, and $i$ is chosen to be sufficiently large to guarantee the pulse reaches equilibrium or degrades due to modulation instability---we use $i = 500$ in our experiments. We take the modulus of the pulse since we are only interested in the evolution of the envelope because of the methods used experimentally. We plot the error in Fig.~\ref{fig:error}. It is clear the parameter space is split into two distinct regions, divided by an incredibly sharp and complex boundary. In the upper-left region the error is $\bigO{1}$; here, the nonlinearity is too strong, and the SPM induces modulation instability. In the lower-right region we find the opposite behaviour---the pulse reaches an equilibrium state and is stable. The increase in dispersion, coupled with modulation, allows the laser to balance the nonlinear effects. The error in this region is on the order of machine precision, or slightly higher due to the discretization.

@@

Also of interest is the energy of the pulse at equilibrium. We compute the energy using \eqref{eq:energy}, or equivalently 
\begin{align}
	E &= \| A_i \|_2^2.
	\label{eq:energy2norm}
\end{align}
Recall from \eqref{eq:order}, this energy is immediately before passing through the output coupler, and so the energy of each pulse emitted by the laser is $(1 - h^2) E$, and $h^2 E$ remains in the cavity. We plot the energy, \eqref{eq:energy2norm}, in the $s$--$b$ plane in Fig.~\ref{fig:energy}. Unsurprisingly, we find the same sharp boundary as we saw in Fig.~\ref{fig:error}. In the unstable region (upper-left) the energy is relatively small. Since the pulse is not sustainable in this region it is not able to foster prominent modes, thus, the intensity is low across the entire pulse. In the stable region (lower-right) we find that the energy smoothly decays as both $b$ and $s$ increase, with the contours being hyperbolic-esque. As $s$ increases, more energy is lost due to modulation since the tails of the pulse become larger. Similarly, as $b$ increases, high frequencies generated by SPM are reduced by our choice of modulation function. From this, we find that a more energetic laser requires weaker dispersion, however, the nonlinearity must be correspondingly low, otherwise modulation instability will destroy the pulse.

@@

Finally, we briefly investigate the effect the order of the components has on the dynamics. In Section~\ref{sec:model}\ref{sec:effects} we justified the order of the gain, nonlinearity, and loss, but, just assumed dispersion preceded modulation. Here, we consider the effect of swapping the dispersion and modulation components. The energy landscape (Fig.~\ref{fig:energyswitch}) is as we would expect. In the previous case of having modulation following dispersion, more energy is removed from the pulse tails. Whereas, having modulation first, more of this energy remains in the cavity. This has the effect of making the laser more powerful than in the previous case, which in turn increases the size of the unstable region. Comparing with Fig.~\ref{fig:energy}, the unstable upper-left region spreads farther rightward, and the energy is larger in the stable region---especially farther away from the origin.

>>>>

\usepackage{subcaption}
\usepackage{tikz}
\usepackage{mathtools}

@@

To proceed we must choose the form of the amplification term, $g(A)$. We assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}. This reduces \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the energy entering the gain fibre is $E_g$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E_g} W \left( \frac{E_g}{\Es} \exp( E_g / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse within the Er-doped fibre. Repeating this process for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 L_D\beta_2/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the output coupler (depending on the layout of the laser cavity the loss may instead take the form $L(A) = R \exp( - \alpha L_T )A$), $L_T$ is the total length of the laser circuit, $L_D$ is the length of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the output coupler, and $\mathcal{F}$ denotes the Fourier transform. Finally, we must consider the modulation. The modulation is considered to be applied externally in which ever way the operator sees fit. For simplicity the representation is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not particularly restrictive. Calcaterra and Boldt~\cite{calcaterra2008a} have shown that a sum of Gaussian functions can approximate any square integrable function. Therefore, our solution extends to any continuous, bounded modulation function.

@@

Revisiting the process maps \eqref{eq:gainform}--\eqref{eq:modform} yields the new, non-dimensional, mappings---after dropping the tildes---
\begin{align}
	G(A) &= \left(E_g^{-1} W \left( a E_g \textrm{e}^{E_g}\right) \right)^{1/2} A, & F(A) &= A \exp( i b |A|^2), \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters, as defined by the values in Table~\ref{tab:values},
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2.
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
Here, $a$ is a measure of the strength of the gain fibre, $h$ is the loss due to absorption and the output coupler, $b$ is the ratio of the frequency shift due to the Kerr nonlinearity to the width of the pulse, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, notice that in this non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:ndparam} have their own independent, non-dimensional parameter.

% Table of parameter values
\begin{table*}[tbp]
	\centering
	\caption{Range of parameter values (taken from~\cite{agrawal2013, burgoyne2014, burgoyneemail, tamura1996, usechak2005}).}
 	\label{tab:values}
 	\begin{tabular}{clll}
		\hline\noalign{\smallskip}
		Symbol & Parameter & Value & Units \\
		\hline\noalign{\smallskip}
		$\alpha$ & Absorption of fibre & $10^{-4}$--$0.3$ & m$^{-1}$ \\
		$\beta_2^f$ & Fibre dispersion & $-50$--$50$ & ps$^2$/km \\
		$\gamma$ & Fibre nonlinearity & $0.001$--$0.01$ & W$^{-1}$ m$^{-1}$ \\
		$\beta_2^g L_D$ & Grating dispersion & $10$--$2000$ & ps$^2$ \\
		$L_T$ & Length of cavity & $10$--$100$ & m \\
		\vspace*{-2mm} $L_f$ & Length of fibre between & $0.15$--$1$ & m \\
		& gain and output coupler & & \\
		$L_g$ & Length of gain fibre & $2$--$3$ & m \\
		$T_M$ & Modulation Time & $15$--$150$ & ps \\
		$R$ & Reflectivity of optical coupler & $0.1$--$0.9$ & -- \\
		$\Es$ & Saturation energy & $10^3$--$10^4$ & pJ \\
		$g_0$ & Small signal gain &  $1$--$10$ & m$^{-1}$ \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table*}

@@

Consider the pulse
\begin{align}
	A = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \textrm{e}^{i \phi_0},
	\label{eq:A0}
\end{align}
where $P$ is the peak power, $C$ is the chirp, $\sigma^2$ is the variance, and $\phi_0$ is the phase (Fig.~\ref{fig:samplegauss}). By computing $\mathcal{L}(A)$ using \eqref{eq:order}, and equating to $A \textrm{e}^{i \phi}$, we obtain the conditions
\begin{align}
	\frac{\sigma^4}{1 - \sigma^2} &= (\sigma^2 + 2 C s^2)^2 + 4 s^4, \label{eq:sigrelation} \\
	\frac{C}{1 - \sigma^2} &= C + 2 \frac{s^2}{\sigma^2} (1 + C^2), \label{eq:chirprelation} \\
	1 &= \frac{W(a E_g \textrm{e}^{E_g})}{E_g} h^2 (1 - \sigma^2)^{1/2} \label{eq:energyrelation}
\end{align}
for \eqref{eq:A0} to be in equilibrium. Manipulating \eqref{eq:sigrelation} and \eqref{eq:chirprelation} we are able to eliminate the chirp, and find an expression in terms of $\sigma$ only, that is,
\begin{align}
	\sigma^8 + 4 s^4 \sigma^6 - 20 s^4 \sigma^4 + 32 s^4 \sigma^2 - 16 s^4 = 0. \label{eq:var}
\end{align}
As \eqref{eq:var} is a quartic in $\sigma^2$, it has an analytic solution, namely,
\begin{equation}
	\begin{split}
		\sigma^2 = \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} & \\
		- s^4 - s^2 & \sqrt{4 + s^4}.
		\label{eq:equilvar}
	\end{split}
\end{equation}
Additionally, we find the chirp to be
\begin{align}
	C = \frac{\sigma^4 \pm \left( \sigma^8 - 16 s^4 \left( 1 - \sigma^2 \right)^2 \right)^{1/2}}{4 s^2 \left( 1 - \sigma^2 \right)},
	\label{eq:chirp}
\end{align}
in terms of $s$ and $\sigma^2$. Moreover, we can obtain the energy of the pulse entering the gain fibre, $E_g$, from \eqref{eq:energyrelation}, and thus, the equilibrium energy at the output coupler. Doing so gives
\begin{align}
	E &= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{1 - h^2 \left( 1 - \sigma^2 \right)^{1/2}}.
	\label{eq:analenergy}
\end{align}
By equating the energy of the pulse in \eqref{eq:A0} and \eqref{eq:analenergy}, the peak power is found to be
\begin{align}
	P &= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{\sqrt{\pi} \sigma \left( 1 - h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}.
	\label{eq:analpower}
\end{align}

The nature of \eqref{eq:var}--\eqref{eq:analpower} makes it difficult to work with. However, by asymptotically expanding \eqref{eq:var}--\eqref{eq:analpower} in terms of the dispersion parameter, $s$, we find the simpler relations
\begin{align}
	\sigma^2 &\sim 2s(1 - s) + \bigO{s^3}, \\
	C &\sim 1 - s + \frac{1}{2}s^2 + \bigO{s^3}, \\
	E &\sim \frac{\log (a h^2)}{1 - h^2} - \frac{1}{1 - h^2} \left( 1 + \frac{h^2 \log(a h^2)}{1 - h^2}  \right) s + \bigO{s^2}, \label{eq:maxenergy} \\
	P &\sim \frac{\log(ah^2)}{\sqrt{2 \pi}(1 - h^2)} s^{-1/2} + \bigO{s^{1/2}},
\end{align}
in the case $s \rightarrow 0$. And,
\begin{align}
	\sigma^2 &\sim 1 - \frac{1}{4}s^{-4} + \frac{3}{8}s^{-8} + \bigO{s^{-12}}, \label{eq:siginfty}\\
	C &\sim \frac{1}{2}s^{-2} - \frac{3}{8}s^{-6} + \bigO{s^{-10}},
\end{align}
in the case $s \rightarrow \infty$. We have no asymptotic expansion of the energy or peak power as $s \rightarrow \infty$ since in this limit the gain is unable to balance the energy lost due to modulation, as indicated by the logarithm in \eqref{eq:analenergy}. Expanding the condition $a h^2 (1 - \sigma^2)^{1/2} > 1$, from \eqref{eq:analenergy}, using \eqref{eq:siginfty} (for a first-order approximation), we find 
\begin{align}
	s^* = \left( \frac{a h^2}{2} \right)^{1/2}
\end{align}
is the maximum value of $s$ to facilitate a sustainable pulse.

@@

We now turn our attention to a more quantitative analysis, and characterize the stability of a laser by exploring the parameter space. We focus on the $s$--$b$ plane in particular since $a$ and $h$ effect only the amplitude of the pulse, whereas the interplay of dispersion and nonlinearity give rise to the instabilities. Additionally, as mentioned, a conservative initial energy will effect the dynamics much less than a more intense pulse since the nonlinearity will be negligible while coming to equilibrium. To identify the stability of a pulse we examine the relative change in the pulse's envelope between consecutive round trips of the cavity. We compute this error by
\begin{align}
	\Delta = \frac{\| |\mathcal{L}^n(A_0)| - |\mathcal{L}^{n-1}(A_0)| \|_2}{\| \mathcal{L}^{n-1}(A_0) \|_2},
	\label{eq:error}
\end{align}
where
\begin{align}
	\| f \|_2^2 \coloneqq \int_{-\infty}^\infty |f|^2 \, \df T,
\end{align}
and $n$ is chosen to be sufficiently large to guarantee the pulse reaches equilibrium or degrades due to modulation instability---we use $n = 500$ in our experiments. We take the modulus of the pulse since we are only interested in the evolution of the envelope because of the methods used experimentally. We plot the error in Fig.~\ref{fig:error}. It is clear the parameter space is split into two distinct regions, divided by an incredibly sharp and complex boundary. In the upper-left region the error is $\bigO{1}$; here, the nonlinearity is too strong, and the SPM induces modulation instability. In the lower-right region we find the opposite behaviour---the pulse reaches an equilibrium state and is stable. The increase in dispersion, coupled with modulation, allows the laser to balance the nonlinear effects. The error in this region is on the order of machine precision, or slightly higher due to the discretization.

@@

Also of interest is the energy of the pulse at equilibrium, which we compute numerically by \eqref{eq:energy}. Recall from \eqref{eq:order}, this energy is immediately before passing through the output coupler, and so the energy of each pulse emitted by the laser is $(1 - h^2) E$, and $h^2 E$ remains in the cavity. We plot the energy in the $s$--$b$ plane in Fig.~\ref{fig:energy}. Unsurprisingly, we find the same sharp boundary as we saw in Fig.~\ref{fig:error}. In the unstable region (upper-left) the energy is relatively small. Since the pulse is not sustainable in this region it is not able to foster prominent modes, thus, the intensity is low across the entire pulse. In the stable region (lower-right) we find that the energy smoothly decays as both $b$ and $s$ increase, with the contours being hyperbolic-esque. As $s$ increases, more energy is lost due to modulation since the tails of the pulse become larger. Similarly, as $b$ increases, high frequencies generated by SPM are reduced by our choice of modulation function. From this, we find that a more energetic laser requires weaker dispersion, however, the nonlinearity must be correspondingly low, otherwise modulation instability will destroy the pulse. Moreover, the maximum energy, which occurs in the limit that both $s$ and $b$ go to zero, is slightly larger than 2.5. With our choice of parameters, \eqref{eq:maxenergy} predicts an energy of 2.55 at the origin, inline with Fig.~\ref{fig:energy}.

@@

Finally, we briefly investigate the effect the order of the components has on the dynamics. In Section~\ref{sec:model}\ref{sec:effects} we justified the order of the gain, nonlinearity, and loss, but, just assumed dispersion preceded modulation. Here, we consider the effect of swapping the dispersion and modulation components. The energy landscape (Fig.~\ref{fig:energyswitch}) is as we would expect. In the previous case of having modulation following dispersion, more energy is removed from the pulse tails. Whereas, having modulation first, more of this energy remains in the cavity. This has the effect of making the laser more powerful than in the previous case, which in turn increases the size of the unstable region. Comparing with Fig.~\ref{fig:energy}, the unstable upper-left region spreads farther rightward, and the energy is larger in the stable region---especially farther away from the origin. Lastly, once again, the energy at the origin is somewhat larger than 2.5, as predicted by \eqref{eq:maxenergy}.
